---
title: "PubH 7462 Assignment 7"
author: "Jeannette Sample"
format: gfm
execute: 
  eval: true
  echo: true
  warning: false
  message: false
editor: visual
---

## Make an ICU status report from the Sepsis dataset.

```{r}
#| echo: false
#| warning: false

library(tidyverse)
library(data.table) ## For the fread function
library(lubridate)

library(tictoc) # For timing
library(googledrive)
library(gt)

source("sepsis_monitor_functions.R")
```

```{r}
#| echo: false
#| warning: false

drive_deauth()
file_link <- "https://drive.google.com/file/d/1pyigdJ2WHT0yVto0CdX0p8R2lZOwG0Tq"
#https://drive.google.com/file/d/1pyigdJ2WHT0yVto0CdX0p8R2lZOwG0Tq/view?usp=share_link

## All data up until now
new_data <- updateData(file_link)

## Include only most recent data
most_recent_data <- new_data %>%
  group_by(PatientID) %>%
  filter(obsTime == max(obsTime))
```

```{r}
#| echo: false
todayTime <- format(Sys.time(),'%A, %B %d, %Y at %H:%M')
```

This report was generated on `r todayTime`.

```{r}
#| echo: false
#df <- makeSepsisDataset()
df <- most_recent_data
```

A table listing the patients who currently have sepsis (if any), along with their most recent heart rate, temperature, and respiratory rate

```{r}
#| echo: false
(tb1 <- most_recent_data %>%
  filter(SepsisLabel == 1) %>%
  group_by(PatientID) %>%
  fill(HR, Temp, Resp) %>%
  slice_tail() %>%
  filter(SepsisLabel == 1) %>%
  ungroup() %>%
  select(PatientID, HR, Temp, Resp) %>%
  gt(rowname_col = "PatientID") %>%
  tab_stubhead(label = "Patient") %>%
  cols_label(HR = "Heart Rate",
             Temp = "Temperature",
             Resp = "Respiratory Rate") %>%
  tab_header(title = "Most recent vitals of patients who currently have sepsis")
)
```

Plots for all patients who have sepsis (if any) that show the complete history of their heart rate, temperature, and respiratory rate during their time in the ICU.

```{r}
#| echo: false
#| warning: false

df1 <- most_recent_data %>%
  filter(SepsisLabel == 1) %>%
  group_by(PatientID) %>%
  slice_tail()

#subset full dataset
df2 <- most_recent_data %>%
  filter(PatientID %in% df1$PatientID)

p <- df2 %>%
    #group_by(PatientID) %>%
    ggplot(aes(x = ICULOS, group = PatientID)) +
    geom_line(aes(y = HR, color = "Heart Rate")) +
    geom_line(aes(y = Temp, color = "Temperature")) + 
    geom_line(aes(y = Resp, color = "Respiratory Rate")) +
    facet_wrap(~PatientID) +
    labs(x = "Hours in ICU",
         y = "Vitals",
         title = "Complete history of vital signs for patients with sepsis",
         color = "Vitals") 

p + theme(legend.position="bottom")
```

A table showing the change in heart rate, temperature, and respiratory rate between the last two measurements for all patients

```{r}
#| echo: false
(tb2 <- most_recent_data %>%
  arrange(PatientID, ICULOS) %>%
  group_by(PatientID) %>%
  fill(HR, Temp, Resp) %>%
  slice_tail(n=2) %>%
# computing difference of each group
  mutate(HR_diff = HR-lag(HR,n=1),
         Temp_diff=Temp-lag(Temp,n=1),
         Resp_diff=Resp-lag(Resp,n=1)) %>%
  slice_tail() %>% 
  select(PatientID, HR_diff, Temp_diff, Resp_diff)  %>%
  ungroup() %>%
  gt(rowname_col = "PatientID") %>%
  tab_stubhead(label = "Patient") %>%
  cols_label(HR_diff = "Heart Rate",
             Temp_diff = "Temperature",
             Resp_diff = "Respiratory Rate") %>%
  tab_spanner(label = "Change in Vitals", columns = c(HR_diff, Temp_diff, Resp_diff)) %>%
  tab_header(title = "Change in vitals between the last two measurements for all patients")
)
```